---
title: "Bicycle-Casestudy"
author: "Mridul"
date: "2023-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

As a junior data analyst, I will perform this case study.I will work for a fictional company , Cyclistic, based in Chicago, USA. Throughout this case study, I follow the six stages of data analysis (ask, prepare, process, analyze, share, and act), and use technical tools (Tableau, R).

# Scenario

I work as a junior data analyst at Cyclistic, a bike-share company in Chicago. The director of marketing believes the company's future success depends on maximizing the number of annual memberships. Therefore,my team wants to understand how casual riders and annual members use Cyclistic bikes differently. From these insights, my team will design a new marketing strategy to convert casual riders into annual members. But first, Cyclistic executives must approve my recommendations, so they must be backed up with compelling data insights and professional data visualizations.

# Ask

Three questions will guide the future marketing program:

-   How do annual members and casual riders use Cyclistic bikes differently?

-   Why would casual riders buy Cyclistic annual memberships?

-   How can Cyclistic use digital media to influence casual riders to become members?

Lily Moreno (the director of marketing and my manager) has assigned you the first question to answer: How do annual members and casual riders use Cyclistic bikes differently?

### Key tasks

1.Identify the business task

-   The main objective is to design marketing strategies aimed at converting casual riders to annual members by understanding how they differ.

2.Consider key stakeholders

-   Director of Marketing (Lily Moreno), Marketing Analytics team, Executive team.

# PREPARE

### Key tasks

1.Download data and store it appropriately.

-   Data has been downloaded and copies have been stored securely on my computer

2.Identify how it's organized.

-   The data is in CSV (comma-separated values) format.

3.Sort and ﬁlter the data.

-   For this analysis, I will be using data for the last 12 months(january2022 - December2022)

4.Determine the credibility of the data.

-   For the purpose of this case study, the datasets are appropriate and will enable me to answer the business questions. All ride ids are unique.

```{r}
#load_package
library(tidyverse)
library(dplyr)
library(skimr)
library(janitor)
library(lubridate)
library(ggplot2)

#data_import

january_2022 <- read_csv("Bicycle/202201-divvy-tripdata.csv")
february_2022 <- read_csv("Bicycle/202202-divvy-tripdata.csv")
march_2022 <- read_csv("Bicycle/202203-divvy-tripdata.csv")
april_2022 <- read_csv("Bicycle/202204-divvy-tripdata.csv")
may_2022 <- read_csv("Bicycle/202205-divvy-tripdata.csv")
june_2022 <- read_csv("Bicycle/202206-divvy-tripdata.csv")
july_2022 <- read_csv("Bicycle/202207-divvy-tripdata.csv")
august_2022 <- read_csv("Bicycle/202208-divvy-tripdata.csv")
september_2022 <- read_csv("Bicycle/202209-divvy-publictripdata.csv")
october_2022 <- read_csv("Bicycle/202210-divvy-tripdata.csv")
november_2022 <- read_csv("Bicycle/202211-divvy-tripdata.csv")
december_2022 <- read_csv("Bicycle/202212-divvy-tripdata.csv")

#bind_data
full_data <- rbind(january_2022,february_2022,march_2022,april_2022,june_2022,july_2022,august_2022,
                   september_2022,october_2022,november_2022,december_2022)


```

# Process

## Key tasks

1.Check the data for errors.

2.Choose your tools.

3.Transform the data so you can work with it effectively.

4.Document the cleaning process.

```{r}
#see_fulldata

head(full_data)
summary(full_data)
colnames(full_data)
str(full_data)
glimpse(full_data)
skim_without_charts(full_data)

#cleaning
get_dupes(full_data)
colSums(is.na(full_data))
final_data <- distinct(full_data) #remove_duplicate
drop_na(final_data)
remove_empty(final_data)
remove_missing(final_data)

cat("Number of rows :",nrow(final_data))
cat("Number of colams :",ncol(final_data))

# adding new columns date, month, day, year and day of the week
final_data$date <- as.Date(final_data$started_at) #The default format is yyyy-mm-dd
final_data$month <- format(as.Date(final_data$date), "%m")
final_data$day <- format(as.Date(final_data$date), "%d")
final_data$year <- format(as.Date(final_data$date), "%Y")
final_data$day_of_week <- format(as.Date(final_data$date), "%A")

# adding new column ride length
final_data$ride_length <- difftime(final_data$ended_at,final_data$started_at) 
is.factor(final_data$ride_length) #checking/verifying ride length data type

#converting ride length data type factor to numeric
final_data$ride_length <- as.numeric(final_data$ride_length)
is.factor(final_data$ride_length)

colnames(final_data)


```
# ANALYZE

## Key tasks

1. Aggregate your data so it’s useful and accessible.

2. Organize and format your data.

3. Perform calculations.

4. Identify trends and relationships.



## Calculate the mean of ride_length and max ride_length

```{r}
final_data %>% 
  group_by(member_casual) %>%
  summarise(average_ride_length = mean(ride_length), median_length = median(ride_length), 
  max_ride_length = max(ride_length), min_ride_length = min(ride_length))
```

## See members and casuals by the total ride taken (ride count)

```{r}
final_data %>% 
  group_by(member_casual) %>% 
  summarize(count=length(ride_id),
            percentage_of_total=(length(ride_id)/nrow(final_data))*100)
```

## Calculate the average ride_length  day_of_the week

```{r}
final_data %>%  
  group_by(member_casual, day_of_week) %>% 
  summarise(average_ride_length = mean(ride_length), .groups="drop")
```

## Calculate the number of rides for users by day_of_week by adding Count of trip_id to Values.

```{r}
final_data %>%
  group_by(day_of_week, member_casual) %>%
  summarise(number_of_rides = n_distinct(ride_id))
```
## see the total rides taken by members and casuals by month 

```{r}
final_data %>%  
  group_by(member_casual, month) %>% 
  summarise(number_of_rides = n(),.groups="drop") %>% 
  arrange(member_casual, month)

```


# SHARE

## Key tasks

1. Determine the best way to share your findings.

2. Create effective data visualizations.

3. Present your findings.

4. Ensure your work is accessible.

<!---#VISUALIZATION-->

## visualize member vs casual distribution

```{r}
final_data %>% 
  group_by(member_casual) %>% 
  summarize(count=length(ride_id),
  percentage_of_total=(length(ride_id)/nrow(final_data))*100) %>%  
  ggplot(aes(x = member_casual, y = count, fill = member_casual)) +
  labs(title ="member vs casual distribution") +
  geom_col(width=0.5, position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```

## visualize the average ride_length  day_of_the week

```{r}
final_data %>%  
  group_by(member_casual, day_of_week) %>% 
  summarise(average_ride_length = mean(ride_length), .groups="drop")%>%
  ggplot(aes(x = day_of_week, y = average_ride_length, fill = member_casual)) +
  geom_col(width=0.5, position = position_dodge(width=0.5)) + 
  labs(title ="Average ride time of Members and Casual riders Vs. Day of the week")
```

## visualize the number of rides for users by day_of_week by adding Count of trip_id to Values.

```{r}
final_data %>%
  group_by(day_of_week, member_casual) %>%
  summarise(number_of_rides = n_distinct(ride_id))%>%  
  ggplot(aes(x = day_of_week, y = number_of_rides, fill = member_casual)) +
  labs(title ="Total rides of Members and Casual riders Vs. Day of the week") +
  geom_col(width=0.5, position = position_dodge(width=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE))
```

## Let's visualize the total rides taken by members and casuals by month 

```{r}
 final_data %>%  
    group_by(member_casual, month) %>% 
    summarise(number_of_rides = n(),.groups="drop") %>% 
    arrange(member_casual, month)  %>% 
    ggplot(aes(x = month, y = number_of_rides, fill = member_casual)) +
    labs(title ="Total rides by Members and Casual riders by Month") +
    theme(axis.text.x = element_text(angle = 45)) +
    geom_col(width=0.5, position = position_dodge(width=0.5)) +
    scale_y_continuous(labels = function(x) format(x, scientific = FALSE))

```


# Act

* Members have more bikes compared to casual riders.

* Casual riders travel for a longer time period.

* June, July, August, September, October are the months with the highest number of riders. Then The number   of casual riders is also good.

* It is decisive that across all the months, the members were more in percent than casuals.

* Weekends are more casual riders.

* Regular rider is higher on weekdays








<!--
## Including Plots

You can also embed plots, for example:

```{r, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
-->

